---
title: "R Notebook"
output: html_notebook
---

let's load in and explore the data. What is our main data again?
```{r}
Bomb_df <- read.csv("/Users/alexandrareich/Desktop/THESIS TIME!/Compilation of GITHUB project code/BOMB_TIME/Data/Hatchery egg condition data 1-11-22_modified.csv")

names(Bomb_df)
library(tidyverse)
#View(Bomb_df)
```

Plot exploratory stuff
Interests: ED(energy density) and LIPID CONTENT vs... eg size? length? Hatch wild?
```{r}
library(ggplot2)
library(cowplot)

#the energy density
#class(Bomb_df$Wild_or_hatch)
ggplot(Bomb_df) + aes(x=Wild_or_hatch, y=BombED.Dry.) + geom_boxplot() +geom_jitter()

#the lipid content
ggplot(Bomb_df) + aes(x=Wild_or_hatch, y=X..lipid) + geom_boxplot() +geom_jitter()

```

05/02/22 addtiion: t-tests!
```{r}
?t.test
library(dplyr)
#H0: not different
#HA: Hatch < Wild

#bomb
names(Bomb_df)
Wild_egg_energy <- Bomb_df %>% 
  filter(Wild_or_hatch == "W") %>%
  select(BombED.Dry.)
    
Hatch_egg_energy <- Bomb_df %>% 
  filter(Wild_or_hatch == "H") %>%
  select(BombED.Dry.)

(T_test_ED <- t.test(Hatch_egg_energy, Wild_egg_energy, alternative="less", var.equal = T) )
##HA: x has a lower mean than y
##assuming homogeneity (constant variance... Is this true tho?)

t.test(Hatch_egg_energy, Wild_egg_energy, alternative="less", var.equal = F) 
##HA: x has a lower mean than y
##assuming homogeneity (Not assuming homogeneity)
##probs not using this one


#lipid
Wild_lipid_energy <- Bomb_df %>% 
  filter(Wild_or_hatch == "W") %>%
  select(X..lipid)
    
Hatch_lipid_energy <- Bomb_df %>% 
  filter(Wild_or_hatch == "H") %>%
  select(X..lipid)

(T_test_lipid <- t.test(Hatch_lipid_energy, Wild_lipid_energy, alternative="less", var.equal = T))


#standard error
T_test_ED$stderr
T_test_lipid$stderr


#I should probably record these t-test values in the resutls section. My results section is a work in progress...
```

Normality tests and such? 05/02/22
```{r}
names(Bomb_df)
mod_E <- aov(BombED.Dry. ~ Wild_or_hatch, data=Bomb_df)
mod_L <- aov(X..lipid ~ Wild_or_hatch, data=Bomb_df)

#mod_E <- aov(BombED.Dry. ~ Wild_or_hatch, data=Bomb_with_covariates_notweird)
#mod_L <- aov(X..lipid ~ Wild_or_hatch, data=Bomb_with_covariates_notweird)

fitted(mod_E)
#plot residuals vs. fitted?
r_E <- resid(mod_E)
ggplot(Bomb_df) + aes(x=r_E, y=mod_E, color=Wild_or_hatch) + geom_point()
plot(resid(mod_E))
ggplot(Bomb_df) + aes(x=r_E) + geom_histogram()
qqnorm(r_E) #concave-down
shapiro.test(r_E) #shit, not normal. Normality of residuals violated. Is normality of residuals an assumption?

plot(fitted(mod_E), r_E)

#ugh! 

fit_L <- fitted(mod_L)
r_L <- resid(mod_L)
qqnorm(r_L)
shapiro.test(r_L) #yes normal, lm ok

plot(r_L) #ok weird, might be an outlier in there

library(onewaytests)
?bf.test  #use this test to test homogeneity resids!!

hist(Bomb_df$BombED.Dry.)
hist(log(Bomb_df$BombED.Dry.))

```


Tangent: 05/02/22
Normality is violated for energy density. Try a glm?
Which glm family to choose??

Or... log trasform? try transforming x first
```{r}
?glm
?box
library(car)
?boxCox
boxCox(mod_E)
#boxCox(mod_E, lambda=seq(-2, 8, 1/5)) #ugh
mod_E_log <- aov(log(BombED.Dry.) ~ Wild_or_hatch, data=Bomb_df)
mod_E_sqrt <- aov(sqrt(BombED.Dry.) ~ Wild_or_hatch, data=Bomb_df)
mod_E_recip <- aov((1/BombED.Dry.) ~ Wild_or_hatch, data=Bomb_df)

plot(resid(mod_E_log))
qqnorm(resid(mod_E_log))
qqnorm(r_E)
qqnorm(resid(mod_E_sqrt)) #none of these look good
qqnorm(resid(mod_E_recip)) #this one looks bad in the other direction

shapiro.test(resid(mod_E_recip))

```


I see more variation in the W for bomb, not so much for lipids. What does this mean?
I need to add in egg size and fish length (and maybe GSI) and see what this means!

DATA WRANGING
Goal: attach bomb data to individual.
What is the indiviudal data called?
(see main R project folder)
Go back into "Cleaned up Analysis and Graphs.rnb and create a clean csv file with the GSI, and save that locally.
```{r}
#GSI
GSI_c <- read.csv("Data/c_GSI_for_boms.csv")

#GSI clean
GSI_c_clean <- read.csv("Data/c_GSI_clean_for_bomb.csv")

#eggs
Eggs_c <- read.csv("Data/c_all_eggs_bomb.csv")
C_other_data <- read.csv("Data/c_all_other_stuff.csv")
  
  
#eggs clean
Eggs_c_clean <- read.csv("Data/c_clean_bomb.csv")

#another idea: add energy content as covariate in egg size analysis?? Uhhh...later. It's bomb time
```

Combine data frames so your bomb dataframe has covariates
```{r}
#joinby
#r_bind
#c_bind
#see data wrangling worksheet

names(Bomb_df)
Bomb_df_relevant <- Bomb_df[, c(2,3,6,7,8,9,10)]

#View(GSI_c_clean)
length(GSI_c_clean$ID)
(Bomb_df$InvestNo)
Bomb_df_relevant$ID<-Bomb_df$InvestNo
(GSI_c$ID)
#join Bomb_df_relevant with GSI_c. This will remove non-egg bearers (there were sone? and then I can remove weird fishies after)
Bomb_with_covariates <- left_join(Bomb_df_relevant, GSI_c, by = "ID" )
length(Bomb_df_relevant$InvestNo)
length(Bomb_with_covariates$ID) #woohoo, it worked

View(Bomb_with_covariates)


```

The subset code addition.
Bomb_data_with_covariates is the dataset to subset from, cause length is attached
```{r}
Bomb_with_covariates <- Bomb_with_covariates %>% filter(Length..mm.>= 559 & Length..mm.<= 608)

Bomb_with_covariates_1 <- Bomb_with_covariates
```


Add in the covariate of average egg size
```{r}
names(Bomb_with_covariates)
#of interest: GSI
#add in: average egg size. This exists somewhere.
#where is my average egg size document?
names(Eggs_c)
names(Eggs_c_clean)
Eggs_c_clean$Fish.ID
#test <- Eggs_c %>%
 # group_by(Fish.ID) %>%
#  select(Fish.ID, Diameter..mm.) %>%
 # mutate(EggDiam = mean(Diameter..mm.))

Eggs_av <- Eggs_c %>%
  group_by(Fish.ID) %>%
  summarize(Eggs_av = mean(Diameter..mm.))

#NICE!
#combine Eggs_av with Bomb_with_covariates
names(Bomb_with_covariates)
Bomb_with_covariates$Fish.ID
names(Eggs_av)

#join them!!
Bomb_with_covariates <- left_join(Bomb_with_covariates, Eggs_av, by = "Fish.ID" )
  
```


Bomb_with_covariates is my dataset right now. Let's check out them covariates!
(and maybe normality assuption too on the energy and lipid data!!)
Pause for now. 04/22/22. Above is what to do next.'

05/02/22
```{r}
Bomb_with_covariates_notweird <- Bomb_with_covariates %>%
  filter(Weird=="n")
Bomb_df <- Bomb_with_covariates_notweird
Bomb_with_covariates_notweird_noNA <- na.omit(Bomb_with_covariates_notweird)


#let's use Bomb_with_covariates_notweird in place of Bomb.Ed.
#Bomb_with_covariates_notweird_noNA <- Bomb_with_covariates
##gets rid of fish number 25, where Milo didn't record the GSI
##ok, that's my DF, the one I want to use.
```

OCTOBER 2022 edits, for the subset:
 t-tests! But bom_df is subsetted
```{r}
?t.test
library(dplyr)
#H0: not different
#HA: Hatch < Wild

#bomb
names(Bomb_df)
Wild_egg_energy <- Bomb_df %>% 
  filter(Wild_or_hatch == "W") %>%
  select(BombED.Dry.)
    
Hatch_egg_energy <- Bomb_df %>% 
  filter(Wild_or_hatch == "H") %>%
  select(BombED.Dry.)

(T_test_ED <- t.test(Hatch_egg_energy, Wild_egg_energy, alternative="less", var.equal = T) )
##HA: x has a lower mean than y
##assuming homogeneity (constant variance... Is this true tho?)

t.test(Hatch_egg_energy, Wild_egg_energy, alternative="less", var.equal = F) 
##HA: x has a lower mean than y
##assuming homogeneity (Not assuming homogeneity)
##probs not using this one


#lipid
Wild_lipid_energy <- Bomb_df %>% 
  filter(Wild_or_hatch == "W") %>%
  select(X..lipid)
    
Hatch_lipid_energy <- Bomb_df %>% 
  filter(Wild_or_hatch == "H") %>%
  select(X..lipid)

(T_test_lipid <- t.test(Hatch_lipid_energy, Wild_lipid_energy, alternative="less", var.equal = T))


#standard error
T_test_ED$stderr
T_test_lipid$stderr


#I should probably record these t-test values in the resutls section. My results section is a work in progress...
```








use this dataframe: Bomb_with_covariates_notweird_noNA
Covariates of interest: 
-Egg size
-Fish length
- wild or hatch

response var:
- Energy denstiy
-Lipid conten

Potentially also:
- GSI
```{r}
names(Bomb_with_covariates_notweird_noNA) #but need to add in 25

#energy
ggplot(Bomb_with_covariates_notweird_noNA) + aes(x=Eggs_av, y=BombED.Dry., color=Wild.or.Hatch) + geom_point()
ggplot(Bomb_with_covariates_notweird_noNA) + aes(x=Length..mm., y=BombED.Dry., color=Wild.or.Hatch) + geom_point()


#lipid
ggplot(Bomb_with_covariates_notweird_noNA) + aes(x=Eggs_av, y=X..lipid, color=Wild.or.Hatch) + geom_point()
ggplot(Bomb_with_covariates_notweird_noNA) + aes(x=Length..mm., y=X..lipid, color=Wild.or.Hatch) + geom_point()

#interactions?
ggplot(Bomb_with_covariates_notweird_noNA) + aes(x=Length..mm., y=Eggs_av, color=Wild.or.Hatch) + geom_point() #ummm

###
hist(Bomb_with_covariates_notweird$BombED.Dry.)
hist(Bomb_with_covariates_notweird$X..lipid)

shapiro.test(Bomb_with_covariates_notweird$BombED.Dry.)
shapiro.test(Bomb_with_covariates_notweird$X..lipid)
```

So... there's just no diff between W and H, no matter how you look at it. Build a model for this? Maybe? Yes?
```{r}
#ED
names(Bomb_with_covariates_notweird_noNA)
mod_universal_E <- aov(BombED.Dry. ~ Wild.or.Hatch + Eggs_av + Length..mm.,Bomb_with_covariates_notweird_noNA)
summary(mod_universal_E)

mod_univ_E_int <- aov(BombED.Dry. ~ Wild.or.Hatch + Eggs_av + Length..mm. + Eggs_av:Length..mm.,Bomb_with_covariates_notweird_noNA)
summary(mod_univ_E_int)

#additional ED explorations
#mod_2_E <- aov(BombED.Dry. ~ Wild.or.Hatch + Eggs_av,Bomb_with_covariates_notweird_noNA)
#summary(mod_2_E) #nothing sig, still

#lipid
mod_universal_L <- aov(X..lipid ~ Wild.or.Hatch + Eggs_av + Length..mm.,Bomb_with_covariates_notweird_noNA)
summary(mod_universal_L)

mod_univ_L_int <- aov(X..lipid ~ Wild.or.Hatch + Eggs_av + Length..mm. + Eggs_av:Length..mm.,Bomb_with_covariates_notweird_noNA)
summary(mod_univ_L_int)

#other 05/24/22
```


Figs for bomb analysis: 
Bomb fig:
```{r}
#energy boxplot
Bomb_df <- Bomb_with_covariates_notweird


#others 
base.e <- ggplot(Bomb_with_covariates_notweird) + aes(x=Eggs_av, y=BombED.Dry., color=Wild.or.Hatch) + geom_point()
ggplot(Bomb_with_covariates_notweird) + aes(x=Length..mm., y=BombED.Dry., color=Wild.or.Hatch) + geom_point()

library(cowplot)
#both boxplots, actually
#class(Bomb_df$Wild_or_hatch)
Bomb<- ggplot(Bomb_df) + aes(x=Wild_or_hatch, y=BombED.Dry.) + geom_boxplot(outlier.alpha = 0) +    geom_jitter(aes(color=Wild_or_hatch)) +
 scale_color_manual(values=c("orange", "blue"), name=NULL)+
  guides(color="none") +
  labs(y= "Energy density (kJ/g)", x= "Fish origin")+
  theme_cowplot()
  
##may want to note that Energy Density units is kJ/g dry weight

#the lipid content
Lipid<-ggplot(Bomb_df) + aes(x=Wild_or_hatch, y=X..lipid) + geom_boxplot(outlier.alpha = 0) +    geom_jitter(aes(color=Wild_or_hatch)) +
 scale_color_manual(values=c("orange", "blue"), name=NULL)+
  labs(y= "Lipid %", x= "Fish origin")+
  theme_cowplot()

library(patchwork)
Bomb + Lipid

#nice plot draft. Probably need to clean up so that Wild and hatchery SAY wild and hatchery, maybe get rid of legend all together


#edit 05/23/22
##making a professional graph. 
##how to do this again??
Bomb2 <- Bomb + coord_cartesian(ylim=c(24,27)) + 
  scale_y_continuous(breaks =c (24, 25, 26, 27), expand=c(0,0)) +
  scale_x_discrete(labels=c("Hatchery", "Wild"))

Lipid2 <- Lipid + guides(color="none") + #get rid of legend
  coord_cartesian(ylim=c(4.2,12)) +
  scale_y_continuous(breaks=c(6, 8, 10,12), expand=c(0,0)) +
  scale_x_discrete(labels=c("Hatchery", "Wild"))

#I like those. Let's patchwork them
Bomb2+Lipid2 #need to combine y-axis??
#how to get rid of x-axis text and then add manual x-axis?
Bomb3 <- Bomb2 + labs(x=element_blank())
Lipid3 <- Lipid2 +labs(x=element_blank())

#p4 <- ggplot(data.frame(l = Bomb2$labels$x, x = 0.3, y = 0.3)) +
  #    geom_text(aes(x, y, label = l)) + 
   #   theme_void() +
    #  coord_cartesian(clip = "off")


#(Bomb2+ Lipid2)/p4
xlab <- Lipid$labels$x
library(grid)      

#Bomb3 + Lipid3
#grid::grid.draw(grid::textGrob(xlab, x = 0.50, y =0.05))      
      
library(gridExtra)
result <- Bomb3 + Lipid3
gt <- patchwork::patchworkGrob(result)
gridExtra::grid.arrange(gt, left = "Disp", bottom = "Hp // Cyl")

result + draw_label(xlab, x=0.5, y=0.5)
#megan code, to save last plot




#try a different way...
grid_bomb <- plot_grid(Bomb3, Lipid3, scale=0.95) #seems to work

#temp_bomb <- plot_grid(NULL, grid_bomb, ncol = 2, rel_widths =c(0.5,6))
bomb_base <- plot_grid(grid_bomb, NULL, ncol = 1, rel_heights = c(9.7,0.3))
bomb_draw <- ggdraw(bomb_base) + draw_label (xlab, x = 0.56, y = 0.07, size = 14) #bomb_draw looks good! let's save that!
#use this size dots or larger tho??


#megan code, to save last plot
#does this work with rmd? let's find out!!
dev.new (width = 3.5, height = 2.225, unit = "in", noRStudioGD = T); last_plot()
ggsave ("Bomb_lipid_compound_graph_subset.jpg", width = dev.size()[1], height = dev.size()[2]); dev.off()
```


Ok so above were my yesterday final graphs
Today: 05/24/22
Let's make sina-violin plot final graphs
I think sina does not allow color, but violin does.
So hatch orange, wild blue
```{r}
library(ggforce)
?geom_sina
?geom_violin

#View(Bomb_df)
#View(Bomb_with_covariates)
length(Bomb_with_covariates$ID)
length(Bomb_df$SIN)
Bomb_df$Weird
#Bomb_with_covariates_notweird

#re-do t-test: Bomb_with_covariates_notwerid

#first make the base: energy
Bomb_v<- ggplot(Bomb_with_covariates_notweird) + aes(x=Wild_or_hatch, y=BombED.Dry.) + geom_violin() +  geom_sina(aes(color=Wild_or_hatch)) +
 scale_color_manual(values=c("orange", "blue"), name=NULL)+
  guides(color="none") +
  labs(y= "Energy density (kJ/g)", x= "Fish origin")+
  theme_cowplot()
#that could work.

#then make the base: lipid
Lipid_v<- ggplot(Bomb_with_covariates_notweird) + aes(x=Wild_or_hatch, y=X..lipid) + geom_violin() +  geom_sina(aes(color=Wild_or_hatch)) +
 scale_color_manual(values=c("orange", "blue"), name=NULL)+
  guides(color="none") +
  labs(y= "Lipid %", x= "Fish origin")+
  theme_cowplot()
#that works
#now lets clean up them graphs
Bomb_v2 <- Bomb_v + coord_cartesian(ylim=c(24,27)) + 
  scale_y_continuous(breaks =c (24, 25, 26, 27), expand=c(0,0)) +
  scale_x_discrete(labels=c("Hatchery", "Wild")) +
  labs(x=element_blank())

Lipid_v2 <- Lipid_v + guides(color="none") + #get rid of legend
  coord_cartesian(ylim=c(4.2,12)) +
  scale_y_continuous(breaks=c(6, 8, 10,12), expand=c(0,0)) +
  scale_x_discrete(labels=c("Hatchery", "Wild")) +
  labs(x=element_blank())

#combine them and save ya' x-label
xlab <- Lipid_v$labels$x

library(gridExtra)
library(patchwork)
#the successful combining plots and labels way
grid_V <- plot_grid(Bomb_v2, Lipid_v2, scale=0.95) #seems to work

#temp_bomb <- plot_grid(NULL, grid_bomb, ncol = 2, rel_widths =c(0.5,6))
V_base <- plot_grid(grid_V, NULL, ncol = 1, rel_heights = c(9.7,0.3))
V_draw <- ggdraw(V_base) + draw_label(xlab, x = 0.56, y = 0.07, size = 14) #bomb_draw looks good! let's save that!
#use this size dots or larger tho??
#LOOKS GREAT! LEt's save this with megans save-a-plot code!
dev.new (width = 3.5, height = 2.225, unit = "in", noRStudioGD = T); last_plot()
ggsave ("Bomb_lipid_VIOLIN_graph.jpg", width = dev.size()[1], height = dev.size()[2]); dev.off()

#aside:
##did I use the correct (non-weird... fish for my egg t-test?? I think most were not weird, we can include WF 25... I think... Coho wild 1 and 2 were weird with only a few eggs) # I think I need to make sure I'm using the not weird ones
##re-do using not weird fish.

##does GSI impact energy or lipid
###nope, we're good there.

```


Lipid fig:
```{r}
#lipid boxplot


#other
base.l <- ggplot(Bomb_with_covariates_notweird_noNA) + aes(x=Eggs_av, y=X..lipid, color=Wild.or.Hatch) + geom_point()
ggplot(Bomb_with_covariates_notweird_noNA) + aes(x=Length..mm., y=X..lipid, color=Wild.or.Hatch) + geom_point()
```



```{r}
summary(mod_E)
plot(mod_E)
```

ON WHICH MODEL ARE MY RESULTS BASED?? 05/23/22
->FROM THE T-TEST!
-> we used the t-test restults because ANOVA showed that covariates were not signficiant... I think
-> did GSI have an effect tho?
```{r}
#ones not used
summary.lm(mod_E) 
summary.lm(mod_universal_E)

#the t-tests
T_test_ED
T_test_lipid

#the standard deviations
##for Energy
sd(Wild_egg_energy$BombED.Dry.)
sd(Hatch_egg_energy$BombED.Dry.)

##for Lipid
sd(Wild_lipid_energy$X..lipid)
sd(Hatch_lipid_energy$X..lipid)


#re-do the t-tests and SD WITHOUT THE WEIRDOS #oh, this is it for the SUBSET results
#Bomb_with_covariates_notweird
Wild_egg_energy_noweird <- Bomb_with_covariates_notweird %>% 
  filter(Wild_or_hatch == "W") %>%
  select(BombED.Dry.)
    
Hatch_egg_energy_noweird <- Bomb_with_covariates_notweird %>% 
  filter(Wild_or_hatch == "H") %>%
  select(BombED.Dry.)

(T_test_ED_noweird <- t.test(Hatch_egg_energy_noweird, Wild_egg_energy_noweird, alternative="less", var.equal = T) )

sd(Hatch_egg_energy_noweird$BombED.Dry.)
sd(Wild_egg_energy_noweird$BombED.Dry.)

#lipid now
Wild_lipid_noweird <- Bomb_with_covariates_notweird %>% 
  filter(Wild_or_hatch == "W") %>%
  select(X..lipid)
    
Hatch_lipid_noweird <- Bomb_with_covariates_notweird %>% 
  filter(Wild_or_hatch == "H") %>%
  select(X..lipid)

(T_test_lipid_noweird <- t.test(Hatch_lipid_noweird, Wild_lipid_noweird, alternative="less", var.equal = T) )

sd(Hatch_lipid_noweird$X..lipid)
sd(Wild_lipid_noweird$X..lipid)

```
GSI sig test -> not sig at all...
```{r}
names(Bomb_with_covariates_notweird_noNA)
universal_bomb <- lm(BombED.Dry. ~ Wild_or_hatch + GSI, data=Bomb_with_covariates_notweird_noNA)
universal_bomb <- lm(BombED.Dry. ~GSI +  Wild_or_hatch, data=Bomb_with_covariates_notweird_noNA)
summary(universal_bomb)
```

