---
title: "Morphcontinued"
author: "Alex"
date: "10/12/2021"
output: html_document
root.dir : "/Users/alexandrareich/Desktop/MORPHO IN ACTION/WORKDIR"
---



First, load the packages

```{r}
library(ggplot2)
library(geomorph)
#library(shapes)
#library(MASS)
#library(ape)
#library(vegan)
#library(rgl)
#library(scatterplot3d)
library(ggfortify)
library(dplyr)
#library(reshape2)
#library(reshape)
library(tidyr)
library(ggforce)
library(Morpho)
library(RColorBrewer)


```


Load my tps file

```{r}
p2.lands <- readland.tps("/Users/alexandrareich/Desktop/THESIS TIME!/Compilation of GITHUB project code/Relaxed_selection_proj_reproducible_science/DATA/pink2021_landmarks.fixed73.deleteextras.appendhere_010522.tps", readcurves=T, specID= "imageID")


#need to order this goddamn chaos above. Need to add in hatch or wild. DO THIS NEXT!!!


p2.lands.unbent.cubic <- readland.tps("/Users/alexandrareich/Desktop/THESIS TIME!/Compilation of GITHUB project code/Relaxed_selection_proj_reproducible_science/DATA/pink2021_cub_010522.tps copy") #no curves, they're all landmarks here. Oh shit, I didn't delete the excess landmakrs. I'll have to go do that. #landmakrs 18 thru 23 need to go because they were just used for unbending.



#I don't think I'll use the quadratic unbending. It did not seem to work as well as the cubic
p2.lands.unbent.quad <- readland.tps("/Users/alexandrareich/Desktop/THESIS TIME!/Compilation of GITHUB project code/Relaxed_selection_proj_reproducible_science/DATA/pink2021_quad_010522.tps copy") #this might be wrong. Too few landmakrs. Maybe I didn't properly update in tpsdig.. Oh wait, it looks ok! Just need to delete the excess landmakrs (18 through 23)



#p2.lands.cub <- p2.lands.unbent.cubic[-18:-23,,] #ok, looks good
  
#p2.lands.quad <- p2.lands.unbent.quad[-18:-23,,]

```


Wild/hatch vector add
```{r}
#wild or hatch vector... 100 fish long...
#p2.lands
#p2.lands[,,20:30]

#FISH 30 IS A DUPLICATE? WTH? W fisj 4 is duplicated? need to check this in TPS DIg in the windows world

#read in the vector
#wildhatch <- read.csv("/Users/alexandrareich/Desktop/MORPHO IN ACTION/pink2021.wildorhatch.andorder.csv")

#wildtest <- wildhatch %>% filter(Wild.or.hatch=="H")
#length(wildtest$Wild.or.hatch) #50 W and 50 H specimens, nice
?duplicated
?anyDuplicated

#anyDuplicated(wildhatch$Fish.ID)
#anyDuplicated(p2.lands$)
#ordered(wildtest$Fish.ID)
#i=1
#for (i in 1:100){
#if(wildhatch$Fish.ID[i]=="W1"){print("yes")}
  #i+1
#}


#i=1
#for (i in 1:100){
#if(wildhatch$Fish.ID[i]!="W"){print(i)}
  #i+1
#}

```



Figuring out the order to see if I can take a shortcut with the wild/hatch labeling...
I actually dont think the shortcut will be legitimate. Might just have to label by hand :(
```{r}
#head(p2.lands)
#head(p2.lands.check.only)

```


Plot all specimens
```{r}
plotAllSpecimens(p2.lands)


```



Add the sliding semi-landmarks to both the original and cubic-unbent dataset.
This will only be done once, and the hash-tagged out.
--Upon some digging it turns out that all the pink semilandmark curve references are the same, so I'll use the same file (curveslide in the MORPHO IN ACTION folder for the pinks).
```{r}
#pinks 21 original
#define.sliders(p2.lands[,,1], nsliders=3)
#needs to be done not in a chunk

#pinks 21 unbent cubic
#define.sliders(p2.lands.unbent.cubic[,,1], nsliders=3)


```

Now transform into Procrustes shape space
```{r}
pinkcurves <- as.matrix(read.csv("/Users/alexandrareich/Desktop/THESIS TIME!/Compilation of GITHUB project code/Relaxed_selection_proj_reproducible_science/DATA/curveslide.csv", header=T))
gpa.p2.orig <- gpagen(p2.lands, curves = pinkcurves)

#gpa.p2.cub <- gpagen(p2.lands.cub, curves = pinkcurves)
#gpa.p2.quad <- gpagen(p2.lands.quad, curves = pinkcurves)

plot(gpa.p2.orig)
#plot(gpa.p2.cub)
#plot(gpa.p2.quad)

#weird with that one long snout. Looks like we may have an outlier here.
```
Now run a PCA
```{r}
pca.p2.orig <- gm.prcomp(gpa.p2.orig$coords)
plot(pca.p2.orig)
#identify(pca.p2.orig$x)


#pca.p2.cub <- gm.prcomp(gpa.p2.cub$coords)
#plot(pca.p2.cub)

#pca.p2.quad <-gm.prcomp(gpa.p2.quad$coords)
#plot(pca.p2.quad)



```






Now what?
Maybe add in the wild/hatchery factor...
How did I do this earlier?
I did this the hard way, there has to be a better way
```{r}
wild.or.hatch <- read.csv("/Users/alexandrareich/Desktop/THESIS TIME!/Compilation of GITHUB project code/Relaxed_selection_proj_reproducible_science/DATA/pink2021.wildorhatch.andorder.csv")
```


Convert to dataframe?
PCA dataframe
```{r}
df.pink2021.orig <- as.data.frame(pca.p2.orig$x[,1:3])
#df.pink2021.cub <- as.data.frame(pca.p2.cub$x[,1:3])
#df.pink2021.quad <- as.data.frame(pca.p2.quad$x[,1:3])

df.pink2021.orig$origin <- wild.or.hatch$Oto
df.pink2021.orig$ID <- wild.or.hatch$Fish.ID

#df.pink2021.cub$origin <- wild.or.hatch$Oto
#df.pink2021.cub$ID <- wild.or.hatch$Fish.ID

#df.pink2021.quad$origin <- wild.or.hatch$Oto
#df.pink2021.quad$ID <- wild.or.hatch$Fish.ID
```

Plot the relative warp analysis(also known as PCA)
```{r}
#original
wildhatch <- c("orange", "blue")

#(plotp<-ggplot(data=df.pink2)+aes(x=Comp1, y=Comp2, color=Wild.or.hatch)+geom_point()+scale_color_manual(values=wildhatch)+
#  stat_ellipse() + theme_cowplot())

ggplot(data=df.pink2021.orig)+aes(x=Comp3, y=Comp1, color=origin)+geom_point()+scale_color_manual(values=wildhatch) + stat_ellipse() + theme_bw() + labs(x= "RW 3", y= "RW 1")

ggplot(data=df.pink2021.orig)+aes(x=Comp2, y=Comp1, color=origin)+geom_point()+scale_color_manual(values=wildhatch) + stat_ellipse() + theme_bw() + labs(x= "RW 2", y= "RW 1")

#cubic unbended
#ggplot(data=df.pink2021.cub)+aes(x=Comp3, y=Comp1, color=origin)+geom_point()+scale_color_manual(values=wildhatch) + stat_ellipse() + theme_bw() + labs(x= "RW 3", y= "RW 1")

#ggplot(data=df.pink2021.cub)+aes(x=Comp2, y=Comp1, color=origin)+geom_point()+scale_color_manual(values=wildhatch) + stat_ellipse() + theme_bw() + labs(x= "RW 2", y= "RW 1")


#quadratic unbended
#ggplot(data=df.pink2021.quad)+aes(x=Comp3, y=Comp1, color=origin)+geom_point()+scale_color_manual(values=wildhatch) + stat_ellipse() + theme_bw() + labs(x= "RW 3", y= "RW 1")

#ggplot(data=df.pink2021.quad)+aes(x=Comp2, y=Comp1, color=origin)+geom_point()+scale_color_manual(values=wildhatch) + stat_ellipse() + theme_bw() + labs(x= "RW 2", y= "RW 1")

#ggplot(data=df.pink2021.quad)+aes(x=Comp2, y=Comp3, color=origin)+geom_point()+scale_color_manual(values=wildhatch) + stat_ellipse() + theme_bw() + labs(x= "RW 2", y= "RW 3")

```

What do these RW's mean???
Original
Comp 1: hump, and a little bit of snout
Comp 2: bendy fish
comp 3: overall fish largeness, some depth involved
Comp4: snout, strech of hump linearly

Cubic unbend
C1: Depth, some snout
C2: bend... I thought I unbent these fish?!?!
C3: Depth again, not snout
C4:snout

Quadratic unbend
C1: hump, snout
C2: bending still... but it is less than other bending?
C3: proportions and a bit of depth
C4:smaller proportions, less significant I think

```{r}
#plotRefToTarget()

#ORIGINAL
#pca plot ref to target
ref <- mshape(gpa.p2.orig$coords) 
ref_p2 <- mshape(gpa.p2.orig$coords) 
#comp1
plotRefToTarget(pca.p2.orig$shapes$shapes.comp1$min, ref, method="vector")
plotRefToTarget(pca.p2.orig$shapes$shapes.comp1$max, ref, method="vector")
#comp2
plotRefToTarget(pca.p2.orig$shapes$shapes.comp2$min, ref, method="vector")
plotRefToTarget(pca.p2.orig$shapes$shapes.comp2$max, ref, method="vector")
#comp3
plotRefToTarget(pca.p2.orig$shapes$shapes.comp3$min, ref, method="vector")
plotRefToTarget(pca.p2.orig$shapes$shapes.comp3$max, ref, method="vector")
#comp4
plotRefToTarget(pca.p2.orig$shapes$shapes.comp4$min, ref, method="vector")
plotRefToTarget(pca.p2.orig$shapes$shapes.comp4$max, ref, method="vector")


#CUBIC
#pca plot ref to target
#ref2 <- mshape(gpa.p2.cub$coords) 
#comp1
#plotRefToTarget(pca.p2.cub$shapes$shapes.comp1$min, ref2, method="vector")
#plotRefToTarget(pca.p2.cub$shapes$shapes.comp1$max, ref2, method="vector")
#comp2
#plotRefToTarget(pca.p2.cub$shapes$shapes.comp2$min, ref2, method="vector")
#plotRefToTarget(pca.p2.cub$shapes$shapes.comp2$max, ref2, method="vector")
#comp3
#plotRefToTarget(pca.p2.cub$shapes$shapes.comp3$min, ref2, method="vector")
#plotRefToTarget(pca.p2.cub$shapes$shapes.comp3$max, ref2, method="vector")
#comp4
#plotRefToTarget(pca.p2.cub$shapes$shapes.comp4$min, ref2, method="vector")
#plotRefToTarget(pca.p2.cub$shapes$shapes.comp4$max, ref2, method="vector")

#QUADRATIC
#pca plot ref to target
#ref3 <- mshape(gpa.p2.quad$coords) 
#comp1
#plotRefToTarget(pca.p2.quad$shapes$shapes.comp1$min, ref3, method="vector")
#plotRefToTarget(pca.p2.quad$shapes$shapes.comp1$max, ref3, method="vector")
#comp2
#plotRefToTarget(pca.p2.quad$shapes$shapes.comp2$min, ref3, method="vector")
#plotRefToTarget(pca.p2.quad$shapes$shapes.comp2$max, ref3, method="vector")
#comp3
#plotRefToTarget(pca.p2.quad$shapes$shapes.comp3$min, ref3, method="vector")
#plotRefToTarget(pca.p2.quad$shapes$shapes.comp3$max, ref3, method="vector")
#comp4
#plotRefToTarget(pca.p2.quad$shapes$shapes.comp4$min, ref3, method="vector")
#plotRefToTarget(pca.p2.quad$shapes$shapes.comp4$max, ref3, method="vector")
```
Plot the wild to hatch comparison: 
```{r}

#plotReftoTarget  ##ORIG
#pinks are in Y.gpa ??
Y.gpa <- gpa.p2.orig
ref4 <- mshape(gpa.p2.orig$coords)
w.p <- numeric()
h.p <- numeric()
i=1

for (i in 1:length(df.pink2021.orig$origin)){
  if(df.pink2021.orig$origin[i]=="W"){
    w.p <- c(w.p, i)
  }else{
    h.p <- c(h.p, i)
  }
}

ref.w.p2 <-mshape(Y.gpa$coords[,,w.p]) #wild mean
ref.h.p2 <-mshape(Y.gpa$coords[,,h.p]) #hatchery mean

#wild to hatch, hatch to wild examine
plotRefToTarget(ref.w.p2, ref.h.p2, method="vector", mag=5) #what is happening here? is wild plotted to hatch or hatch plotted to wild?
plotRefToTarget(ref.h.p2, ref.w.p2, method="vector", mag=5) #hatchery dots, wild arrows
#HYPOTHESIS SUPPORTED! DAMN!


plotRefToTarget(ref.w.p2, ref.h.p2, method="vector", mag=5) #wild dot, hatchery arrows
plotRefToTarget(ref.w.p2, ref4, method="vector", mag=5) #wild to mean fish
plotRefToTarget(ref.h.p2, ref4, method="vector", mag=5) #hatchery to mean fish
```

Quadratioc morpho vidualization comparison:
```{r}
#plotReftoTarget  ##ORIG
#pinks are in Y.gpa ??
#Y.gpa <- gpa.p2.quad
#ref4 <- mshape(gpa.p2.quad$coords)
#w.p <- numeric()
#h.p <- numeric()
#i=1

#for (i in 1:length(df.pink2021.orig$origin)){
  #if(df.pink2021.orig$origin[i]=="W"){
   # w.p <- c(w.p, i)
  #}else{
   # h.p <- c(h.p, i)
  #}
#}

#ref.w.p2 <-mshape(Y.gpa$coords[,,w.p]) #wild mean
#ref.h.p2 <-mshape(Y.gpa$coords[,,h.p]) #hatchery mean

#wild to hatch, hatch to wild examine
#plotRefToTarget(ref.w.p2, ref.h.p2, method="vector", mag=5) #what is happening here? is wild plotted to hatch or hatch plotted to wild?
#plotRefToTarget(ref.h.p2, ref.w.p2, method="vector", mag=5) #hatchery dots, wild arrows
#HYPOTHESIS SUPPORTED! DAMN!


#plotRefToTarget(ref.w.p2, ref.h.p2, method="vector", mag=5) #wild dot, hatchery arrows
#plotRefToTarget(ref.w.p2, ref4, method="vector", mag=5) #wild to mean fish
#plotRefToTarget(ref.h.p2, ref4, method="vector", mag=5) #hatchery to mean fish
```



PCA influence
Noted: bending (RW2) is decreased with the quadratic unbending(desirable?) and increased with cubic unbending (undesirable?).... I should read more about unbending...
```{r}
#ORIGINAL
#can you also plot the ID of the important PCs???
pvar2 <- (pca.p2.orig$sdev^2)/(sum(pca.p2.orig$sdev^2)) #sooo...does this work?
names(pvar2) <- seq(1:length(pvar2))
barplot(pvar2, xlab= "Principal Components", ylab = "% Variance")

#also plot the CVA, please
vari <- gpa.p2.orig$coords
facto <- df.pink2021.orig$origin

CVA.pink <- CVA(vari, groups=facto)


#CUBIC UNBEND
#pvar <- (pca.p2.cub$sdev^2)/(sum(pca.p2.cub$sdev^2)) #sooo...does this work?
#names(pvar) <- seq(1:length(pvar))
#barplot(pvar, xlab= "Principal Components", ylab = "% Variance")

#also plot the CVA, please
#vari <- gpa.p2.cub$coords
#facto <- df.pink2021.cub$origin

#CVA.pink <- CVA(vari, groups=facto)


#QUADRATIC UNBEND
#pvar <- (pca.p2.quad$sdev^2)/(sum(pca.p2.quad$sdev^2)) #sooo...does this work?
#names(pvar) <- seq(1:length(pvar))
#barplot(pvar, xlab= "Principal Components", ylab = "% Variance")

#also plot the CVA, please
#vari <- gpa.p2.quad$coords
#facto <- df.pink2021.quad$origin

#CVA.pink <- CVA(vari, groups=facto)


```


Significance test?
MANCOVA
```{r}
#ORIGINAL
Y.gpa <- gpa.p2.orig
gdf <- geomorph.data.frame(Y.gpa, origin = df.pink2021.orig$origin)

fit.p <- procD.lm(coords ~ log(Csize) + origin, data = gdf,  #all coords taken into account
iter = 9999, print.progress = FALSE, RRPP=TRUE)
summary(fit.p)  #says YES, significant difference between wild and hatch

#are interaction effects a thing?
fit.p.int <- procD.lm(coords ~ log(Csize) + origin + log(Csize)*origin, data = gdf,  #all coords taken into account
iter = 9999, print.progress = FALSE, RRPP=TRUE)
summary(fit.p.int)  #interaction effects not significant!

#Maybe test the relative warps


#CUBIC
#Y.gpa <- gpa.p2.cub
#gdf <- geomorph.data.frame(Y.gpa, origin = df.pink2021.cub$origin)

#fit.p <- procD.lm(coords ~ log(Csize) + origin, data = gdf,  #all coords taken into account
#iter = 9999, print.progress = FALSE, RRPP=TRUE)
#summary(fit.p)  #says YES, significant difference between wild and hatch. Centroid size is less sig than original

#are interaction effects a thing?
#fit.p.int <- procD.lm(coords ~ log(Csize) + origin + log(Csize)*origin, data = gdf,  #all coords taken into account
#iter = 9999, print.progress = FALSE, RRPP=TRUE)
#summary(fit.p.int)  #interaction effects not significant!

#QUADRATIC
#Y.gpa <- gpa.p2.quad
#gdf <- geomorph.data.frame(Y.gpa, origin = df.pink2021.quad$origin)

#fit.p <- procD.lm(coords ~ log(Csize) + origin, data = gdf,  #all coords taken into account
#iter = 9999, print.progress = FALSE, RRPP=TRUE)
#summary(fit.p)  #says YES, significant difference between wild and hatch. Centroid size is less sig than original

#are interaction effects a thing?
#fit.p.int <- procD.lm(coords ~ log(Csize) + origin + log(Csize)*origin, data = gdf,  #all coords taken into account
#iter = 9999, print.progress = FALSE, RRPP=TRUE)
#summary(fit.p.int)  #interaction effects not significant!

#check resids or anything like that?



#may 05/03/22
##the original, without the date
Y.gpa <- gpa.p2.orig
gdf <- geomorph.data.frame(Y.gpa, origin = df.pink2021.orig$origin)

fit.p <- procD.lm(coords ~ log(Csize) + origin, data = gdf,  #all coords taken into account
iter = 9999, print.progress = FALSE, RRPP=TRUE)
summary(fit.p)  

#add date. Is date sig? 
##match up date in here!
##add date just like we add origin!!
##date:
#use dplyr magic to combine date from dataframe:
#######
#With
#dataframe: 
#df.pink2021.orig

#what dataframe is needed for this combination??
p_2_old <- read.csv("/Users/alexandrareich/Desktop/THESIS TIME!/Compilation of GITHUB project code/Relaxed_selection_proj_reproducible_science/DATA/Male.p2.Rdata.2_alt_ID.csv") #not working...why?
names(p_2_old)
p_2_old$ID
df.pink2021.orig$ID
##sigh. They're not identical. I'll need to data wrangle - 05/03/22
df_pink2021_orig_05 <- left_join(df.pink2021.orig, p_2_old, by="ID")
names(df_pink2021_orig_05)
df_pink2021_orig_05$Date


#05/27/22: I'll make a Jdate adjustment, it will make things work betteer
library(lubridate)
names(df_pink2021_orig_05)
Dates_toJ <- mdy(df_pink2021_orig_05$Date)
Julian_morph <- julian(Dates_toJ, origin=as.Date("2021-01-01"))
class(Julian_morph)
str(Julian_morph)
df_pink2021_orig_05$Julian <- as.vector(Julian_morph)


Y.gpa <- gpa.p2.orig
gdf <- geomorph.data.frame(Y.gpa, origin = df_pink2021_orig_05$origin, date=df_pink2021_orig_05$Julian )
gdf_Julian <- gdf
df_p2_reversed_Julian <- as.data.frame(gdf$coords, gdf$origin, gdf$date, gdf$Csize)
names(gdf)

fit.p_04 <- procD.lm(coords ~ log(Csize) + origin, data = gdf,  #all coords taken into account
iter = 9999, print.progress = FALSE, RRPP=TRUE)
summary(fit.p_04) 

fit.p_05 <- procD.lm(coords ~ log(Csize) + origin + date, data = gdf,  #all coords taken into account
iter = 9999, print.progress = FALSE, RRPP=TRUE)

fit.p_06 <- procD.lm(coords ~ log(Csize) + date + origin, data = gdf,  #all coords taken into account
iter = 9999, print.progress = FALSE, RRPP=TRUE)

fit.p_07 <- procD.lm(coords ~ log(Csize) +origin+ date, data = gdf,  #all coords taken into account
iter = 9999, print.progress = FALSE, RRPP=TRUE)

fit.p_08 <- procD.lm(coords ~ log(Csize) * date * origin, data = gdf,  #all coords taken into account
iter = 9999, print.progress = FALSE, RRPP=TRUE) #no interactions

#in the case of date not sig, we go back to original


summary(fit.p_04)#this one if date is not sig 
summary(fit.p_05)
summary(fit.p_06) #this one? #I think this one... #with origin last...?
summary(fit.p_07)
summary(fit.p_08)

#date is significant predictor of shape, until I tested J-date, now its not.


#gotta graph to figure out date relationship on coords

#AIC(fit.p_04)
#AIC(fit.p_05)
anova(fit.p_04, fit.p_05) #fit 5 is better than 4. Keep the mod with date. Graph shape vs. date? How?
names(fit.p_05)
fit.p_05$aov.table


  


```


#sep out wild hatch


######################
##OTHER morpho analyses

Where's my OG pink 2021 morpho data?
THESIS TIME!/Field Data Year 2/FISH
```{r}
p2.males <- read.csv("/Users/alexandrareich/Desktop/THESIS TIME!/Compilation of GITHUB project code/Relaxed_selection_proj_reproducible_science/DATA/Male.p2.Rdata.2.csv")
```


Graph that
```{r}
#use graphs you already made, dummy!
#fuck, can't find them
names(p2.males)
hist(p2.males$Body.depth.mm.)
hist(p2.males$Snout.length.mm.)
hist(p2.males$Length.mm.)
ggplot(p2.males) + aes(x=Length.mm., y=Body.depth.mm.) + geom_point() 
ggplot(p2.males) + aes(x=Length.mm., y=Snout.length.mm.) + geom_point()

#additions: 05/03/22
library(ggpubr)
ggqqplot(p2.males$Snout.length.mm.)
ggqqplot(log(p2.males$Snout.length.mm.))

ggqqplot(p2.males$Body.depth.mm.)
ggqqplot(log(p2.males$Body.depth.mm.)) #log that!1

ggqqplot(p2.males$Length.mm.)
```
Better graphs
```{r}
library(cowplot)

p2.males1 <- p2.males %>% filter(Otolith.reading!= "Overground")

ggplot(p2.males1) + aes(x=Length.mm., y=Body.depth.mm., color=Otolith.reading) + geom_point(size=3) +  scale_color_manual(values=c("blue","orange")) + theme_cowplot()+
  guides(color= "none")+
  geom_smooth(method = "lm")
  #theme(text=element_text(family="Times New Roman"))+
  #guides(color= "none")#+
  #labs()
  #theme(legend.position = c(0.03, 0.92))+theme(legend.title = element_blank(), legend.box.background = element_rect(colour = "black", fill="white", linetype="solid"))+





ggplot(p2.males1) + aes(x=Length.mm., y=Snout.length.mm., color=Otolith.reading) + geom_point(size = 3) +
  scale_color_manual(values=c("blue","orange")) + theme_cowplot()+
  guides(color= "none") +
  geom_smooth(method = "lm")
#theme(text=element_text(family="Times New Roman"))+

#looks like WILD > HATCHERY, hypothesis supported. Now let's test the hypothesis


#05/06/22:
#logging them graphs
#logged p2 snout
ggplot(p2.males1) + aes(x=Length.mm., y=log(Snout.length.mm.), color=Otolith.reading) + geom_point(size = 3) +
  scale_color_manual(values=c("blue","orange")) + theme_cowplot()+
  guides(color= "none") +
  geom_smooth(method = "lm") +
  labs(x= "MEHP length (mm)", y= "log(Snout (mm))")

#logged p2 depth
ggplot(p2.males1) + aes(x=Length.mm., y=log(Body.depth.mm.), color=Otolith.reading) + geom_point(size=3) +  scale_color_manual(values=c("blue","orange")) + theme_cowplot()+
  guides(color= "none")+
  geom_smooth(method = "lm") +
  labs(x= "MEHP length (mm)", y= "log(Depth (mm))")

#NEED TO FIX THE DAMN AXES AND IM NOT DOING IT RIGHT NOW!

#05/26/22
##moving the graphs over to work on them more: to simply graphing

#08/16/22 explore
#logged p2 depth and logged length
ggplot(p2.males1) + aes(x=log(Length.mm.), y=log(Body.depth.mm.), color=Otolith.reading) + geom_point(size=3) +  scale_color_manual(values=c("blue","orange")) + theme_cowplot()+
  guides(color= "none")+
  geom_smooth(method = "lm") +
  labs(x= "MEHP length (mm)", y= "log(Depth (mm))")
##MORE COMPLICATED BUT NOT WORTH IT.
```
which were strays (for building the csv)
```{r}
library(dplyr)
strays <- p2.males1 %>% filter(Location != "Armstrong")
strays <- strays %>% filter(Otolith.reading == "PORT ARMSTRONG")
#identifies the male strats
```


Hypothesis test- updated 05/04/22
```{r}
mod.p2021.depth <- lm(Body.depth.mm. ~ Length.mm. + Otolith.reading,data=p2.males1)
mod.p2021.depth_log <- lm(log(Body.depth.mm.) ~ Length.mm. + Otolith.reading,data=p2.males1)
names(p2.males1)



ggplot(p2.males1) + aes(x=Date, y=Body.depth.mm.) + geom_boxplot() + geom_jitter(aes(color=Otolith.reading, shape=Location))
ggplot(p2.males1) + aes(x=Date, y=Snout.length.mm.) + geom_boxplot() + geom_jitter(aes(color=Otolith.reading, shape=Location))


mod.p2021.snout <- lm(Snout.length.mm. ~ Length.mm. + Otolith.reading,data=p2.males1)
mod.p2021.snout_log <- lm(log(Snout.length.mm.) ~ Length.mm. + Otolith.reading,data=p2.males1)

summary(mod.p2021.depth_log) #the important part for analysis
summary(mod.p2021.snout_log)

#finding them means 05/09/22
w<-p2.males1 %>% filter(Otolith.reading=="No Mark") 
mean(w$Body.depth.mm.) #mean of w male depth for p2

h<-p2.males1 %>% filter(Otolith.reading=="PORT ARMSTRONG") 
mean(h$Body.depth.mm.) #mean of h male depth for p2

sd(w$Body.depth.mm.) #sd's, for my table
sd(h$Body.depth.mm.)
 
mean(w$Snout.length.mm.) #mean of w snout for p2

mean(h$Snout.length.mm.)  #mean of hatch snout for p2

sd(w$Snout.length.mm.)#standard deviations, for my table, for snout lenght
sd(h$Snout.length.mm.)

notresults1 <- t.test(log(h$Snout.length.mm.), log(w$Snout.length.mm.))
notresults2 <- t.test(h$Snout.length.mm., w$Snout.length.mm., alternative = "less")
notresutls3 <- t.test(h$Body.depth.mm., w$Body.depth.mm., alternative = "less")
notresults4 <- t.test(h$Body.depth.mm., w$Body.depth.mm.)

#write.csv(p2.males1, "p2.males1.csv")
#idea: reclassify dates?
##as early, medium, late dates?
#p2_male_date_adj <- p2.males1 %>% 
 # mutate(
#  Date_recat = if(Date== "8/25/21"|| "8/27/21"||"8/31/21"){
 #   Date_recat = "early"
#  } else {
 #   if(Date== "9/1/21"|| "9/3/21"){
  #    Date_recat = "middle"
   # } else {
    #  if(Date== "9/9/21"|| "9/10/21"|| "9/13/21" ){
     #   Date_recat = "late"
    #  }
    #}
  #}
#)


#FUCK. Going to have to make another csv for this...
#lets do this later

#not working
#i<-1
#Date_recat <- numeric(0)
#p2_male_date_adj_2 <- for (i in 1:length(p2.males1$Date)){
 # Date_recat[i] = if(p2.males1$Date[i]== "8/25/21"|| "8/27/21"||"8/31/21"){
#    Date_recat = "early"
 # } else {
  #  if(p2.males1$Date[i]== "9/1/21"|| "9/3/21"){
   #   Date_recat[i] = "middle"
  #  } else {
   #   if(p2.males1$Date[i]== "9/9/21"|| "9/10/21"|| "9/13/21" ){
    #    Date_recat[i] = "late"
     # }
#    }
 # }
#}

#View(p2_male_date_adj)

#ok, fixed the date bullshit
p2.males_dateadj<- read.csv("/Users/alexandrareich/Desktop/THESIS TIME!/Compilation of GITHUB project code/Relaxed_selection_proj_reproducible_science/DATA/p2.males2_dateadj.csv")

#GRAPHS WITH DATE
ggplot(p2.males_dateadj) + aes(x=Date_adj, y=Body.depth.mm.) + geom_boxplot() + geom_jitter(aes(color=Otolith.reading, shape=Location))
ggplot(p2.males_dateadj) + aes(x=Date_adj, y=Snout.length.mm.) + geom_boxplot() + geom_jitter(aes(color=Otolith.reading, shape=Location))

#MODELS WITH DATE
Snout_global_pinkodd <- lm(log(Snout.length.mm.) ~ Length.mm. + Otolith.reading +  Location + Date_adj, data=p2.males_dateadj)
summary(Snout_global_pinkodd)
  
  
Depth_global_pinkodd <- lm(log(Body.depth.mm.) ~ Length.mm. + Otolith.reading +  Location + Date_adj, data=p2.males_dateadj)
summary(Depth_global_pinkodd) #date not sig to this model, but lcoaiton is

Depth_global_pinkodd_int <- lm(log(Body.depth.mm.) ~ Length.mm. + Otolith.reading +  Location + Date_adj + Length.mm.:Otolith.reading + Length.mm.:Location + Length.mm.:Date_adj + Otolith.reading:Location + Otolith.reading:Date_adj + Location:Date_adj + Length.mm.:Otolith.reading:Location, data=p2.males_dateadj)
summary(Depth_global_pinkodd_int)

########################3
#Date edits 05/26/22
##I'll use lubridate package to test for julian date as a continuous variable
library(dplyr)
library(lubridate)
Julian_pinkodd <- julian(mdy(p2.males_dateadj$Date), origin=as.Date("2021-01-01") )
Julian_pinkodd
p2.males_dateadj$Julian <- Julian_pinkodd

#nice. now test for Julian date:
Snout_global_pinkodd_julian <- lm(log(Snout.length.mm.) ~ Length.mm. + Otolith.reading + Julian, data=p2.males_dateadj)
summary(Snout_global_pinkodd_julian) #ok well, that's important. DAte IS signficant here, barely
  
Depth_global_pinkodd_julian <- lm(log(Body.depth.mm.) ~ Length.mm. + Otolith.reading + Julian, data=p2.males_dateadj)
summary(Depth_global_pinkodd_julian) 

Depth_global_pinkodd_julian_check <- lm(log(Body.depth.mm.) ~ Length.mm. + Julian + Otolith.reading, data=p2.males_dateadj)
summary(Depth_global_pinkodd_julian_check)

Depth_global_pinkodd_julian_global <- lm(log(Body.depth.mm.) ~ Length.mm. * Julian * Otolith.reading, data=p2.males_dateadj)
summary(Depth_global_pinkodd_julian_global)

names(p2.males_dateadj)
#View(p2.males_dateadj)




#how does difference in date look for snout and hump in pink 2021?
#EXPLORATORY DATE PLOTS
ggplot(p2.males_dateadj) + aes(x=Julian, y=Body.depth.mm.) +geom_point(aes(color=Location, shape=Otolith.reading)) +
  geom_smooth()
ggplot(p2.males_dateadj) + aes(x=Julian, y=Snout.length.mm.) +geom_point(aes(color=Location, shape=Otolith.reading)) +
  geom_smooth()


```


05/26/22
Finally testing things without the snakey fish in the pink 2021 dataset
```{r}


#not ripe fish remodved test
length(p2.males_dateadj$Length.mm.) #98
Ripe <- p2.males_dateadj %>% filter(Length.mm.>80)
length(Ripe)
#Lovers, Sashin as blocking variables? How to test that?
#not ripe fish remodved test
length(p2.males_dateadj$Length.mm.) #98
Ripe <- p2.males_dateadj %>% filter(Body.depth.mm.>80)
length(Ripe$Length.mm.)   #94. Perfect. 4 questionable fish removed

Depth_global_pinkodd_julian_ripe <- lm(log(Body.depth.mm.) ~ Length.mm. + Otolith.reading + Julian, data=Ripe)
summary(Depth_global_pinkodd_julian_ripe) #no snakey
summary(Depth_global_pinkodd_julian) #yes snakey
#same signficances, but more intesne without snakey fish


```


05/19/22: Length mean of pink 2021 (odd-year pink)
Whats my dataframe??
```{r}
names(p2.males_dateadj)

p2021_H <- p2.males_dateadj %>% filter(Otolith.reading=="PORT ARMSTRONG")
p2021_W <-p2.males_dateadj %>% filter(Otolith.reading=="No Mark")

mean(p2021_H$Length.mm.)
mean(p2021_W$Length.mm.)


#ok, let's do SD
#pink 2021 sd
sd(p2021_H$Length.mm.)
sd(p2021_W$Length.mm.)

#t-test of this
t.test(p2021_H$Length.mm., p2021_W$Length.mm., alternative = "two.sided", var.equal = T)

```


02/11/22
Test the lovers vs sashin issue on p2.
-sahin had snagging (some), lovers did not.
```{r}
#wheres my dataframe?
#hump
names(p2.males1)
ggplot(p2.males1) + aes(y=Body.depth.mm., x= Location, color=Otolith.reading) +
  geom_boxplot() +geom_jitter()
ggplot(p2.males1) + aes(y=Body.depth.mm., x= Location) +
  geom_boxplot() +geom_jitter()  #oof, looks to be a difference here
ggplot(p2.males1) +aes(y=Body.depth.mm., color= Location, x=Length.mm.) +geom_point() +geom_smooth(method = "lm")


#snout
ggplot(p2.males1) + aes(y=Snout.length.mm., x= Location, color=Otolith.reading) +
  geom_boxplot() +geom_jitter()
ggplot(p2.males1) + aes(y=Snout.length.mm., x= Location) +
  geom_boxplot() +geom_jitter()  #oof, looks to be a difference here
ggplot(p2.males1) +aes(y=Snout.length.mm., color= Location, x=Length.mm.) +geom_point() +geom_smooth(method = "lm")



#formal test
summary.lm(mod.p2021.depth)
justwild.p2 <- p2.males1 %>% filter(Otolith.reading == "No Mark")
mod.lov.depth <- aov(Body.depth.mm. ~ Length.mm. + Location, data=justwild.p2)
summary.lm(mod.lov.depth)
```





Well, it looks like there was a bias in lovers cove sampling. OR fish are just bigger from lovers cove.



05/27/22
MORPHO DATE!! : fitp_04 is the way to go (Julian date (numeric) not sig...)
geodata frame:gdf_Julian; Ygpa_Julian; df_p2_reversed_Julian ; df_pink2021_orig_05
```{r}
#LETS GRAPH IT ANYWAY! 
library(viridis)
#ggplot(df_pink2021_orig_05) + aes(x=Comp1, y=Comp3, color=Julian) +
  #geom_point()+
  #scale

#need mean consensus shape per Julian date

unique(gdf_Julian$date)
#coords_252 <- df_p2_reversed_Julian  %>% filter(date==252)
coords_array <- gdf_Julian$coords
coords_df <- as.data.frame(coords_array)

#i<-1
#sf_252=numeric()
#for(i in 1:length(gdf_Julian$date)){
 # if(gdf_Julian$date==252){
  #  df_252[i]= 
  #}
  #i+1
#}


day_252 <- numeric()
day_251 <- numeric()
day_255 <- numeric()
day_238 <- numeric()
day_245 <- numeric()
day_242 <- numeric()
day_236 <- numeric()
day_243 <- numeric()

#Jdate_calc <- function(day_number, day_label){
i=1
for(i in 1:length(gdf_Julian$date)){
  if(df_pink2021_orig_05$Julian[i]==252){
    day_252 <- c(day_252, i)
  }
}
#}

i=1
for(i in 1:length(gdf_Julian$date)){
  if(df_pink2021_orig_05$Julian[i]==251){
    day_251 <- c(day_251, i)
  }
}


i=1
for(i in 1:length(gdf_Julian$date)){
  if(df_pink2021_orig_05$Julian[i]==255){
    day_255 <- c(day_255, i)
  }
}

i=1
for(i in 1:length(gdf_Julian$date)){
  if(df_pink2021_orig_05$Julian[i]==238){
    day_238 <- c(day_238, i)
  }
}


i=1
for(i in 1:length(gdf_Julian$date)){
  if(df_pink2021_orig_05$Julian[i]==245){
    day_245 <- c(day_245, i)
  }
}

i=1
for(i in 1:length(gdf_Julian$date)){
  if(df_pink2021_orig_05$Julian[i]==242){
    day_242 <- c(day_242, i)
  }
}


i=1
for(i in 1:length(gdf_Julian$date)){
  if(df_pink2021_orig_05$Julian[i]==236){
    day_236 <- c(day_236, i)
  }
}

i=1
for(i in 1:length(gdf_Julian$date)){
  if(df_pink2021_orig_05$Julian[i]==243){
    day_243 <- c(day_243, i)
  }
}


ref_252 <-mshape(gdf_Julian$coords[,,day_252])
ref_251 <-mshape(gdf_Julian$coords[,,day_251])
ref_255 <-mshape(gdf_Julian$coords[,,day_255])
ref_238 <-mshape(gdf_Julian$coords[,,day_238])
ref_245 <-mshape(gdf_Julian$coords[,,day_245])
ref_242 <-mshape(gdf_Julian$coords[,,day_242])
ref_236 <-mshape(gdf_Julian$coords[,,day_236])
ref_243 <-mshape(gdf_Julian$coords[,,day_243])

#class(ref_243)
#data.frame(ref_243)

df_252 <-data.frame(ref_252)
df_251 <-data.frame(ref_251)
df_255 <-data.frame(ref_255)
df_238 <-data.frame(ref_238)
df_245 <-data.frame(ref_245)
df_242 <-data.frame(ref_242)
df_236 <-data.frame(ref_236)
df_243 <-data.frame(ref_243)


pallete <- scale_color_grey()
#ggplot() + aes() + geom_point(data=df_236,aes(x=X, y=Y)) +
  #geom_point(data=df_238,aes(x=X, y=Y), color=grey12.5)+
  #geom_point(data=df_242,aes(x=X, y=Y), color=grey15) +
  #geom_point(data=df_243,aes(x=X, y=Y)) +
  #geom_point(data=df_245,aes(x=X, y=Y))+
  #geom_point(data=df_251,aes(x=X, y=Y))+
  #geom_point(data=df_252,aes(x=X, y=Y))+
  #geom_point(data=df_255,aes(x=X, y=Y))
#BELOW NEEDS EDIT, use for example

#canprobs combine into one df, with date
df_252$Jday <-252
df_251$Jday  <-251
df_255$Jday  <-255
df_238$Jday  <-238
df_245$Jday  <-245
df_242$Jday  <-242
df_236$Jday  <-236
df_243$Jday  <-243

#NOW COMBINE THE DATAFRAMES.. STACK THEM!!
?rbind
?bind_rows
?rbind.fill
?vec_is_list
df_Jdate_master <-bind_rows(tibble(X=df_236$X, Y=df_236$Y, Jday=df_236$Jday), 
                            tibble(X=df_238$X, Y=df_238$Y, Jday=df_238$Jday), 
                            tibble(X=df_242$X, Y=df_242$Y, Jday=df_242$Jday), 
                            tibble(X=df_243$X, Y=df_243$Y, Jday=df_243$Jday),
                            tibble(X=df_245$X, Y=df_245$Y, Jday=df_245$Jday),
                            tibble(X=df_251$X, Y=df_251$Y, Jday=df_251$Jday),
                            tibble(X=df_252$X, Y=df_252$Y, Jday=df_252$Jday), 
                            tibble(X=df_255$X, Y=df_255$Y, Jday=df_255$Jday)
                            )
#nice!

#NOW plot
library(viridis)
ggplot(df_Jdate_master) + aes(x=X, y=Y, color=as.factor(Jday)) +
  geom_point()+
  scale_color_viridis_d() +
  theme_void()

ggplot(df_Jdate_master) + aes(x=X, y=Y, color=as.factor(Jday)) +
  geom_point()+
  scale_color_grey()+
  theme_void()


#ggsave
dev.new (width = 9, height = 3, unit = "in", noRStudioGD = T); last_plot()
#ggsave ("Morpho_date_grey_draft.jpg", width = dev.size()[1], height = dev.size()[2]); dev.off()


#BELOW NEEDS EDIT, use for example
#wild to hatch, hatch to wild examine
#plotRefToTarget(ref.w.p2, ref.h.p2, method="vector", mag=5) #what is happening here? is wild plotted to hatch or hatch plotted to wild?
#plotRefToTarget(ref.h.p2, ref.w.p2, method="vector", mag=5) #hatchery dots, wild arrows


```




No snaky fish tests 06/02/22
```{r}
p2.males1
#ggplot(p2.males1) + aes(y=Body.depth.mm., x=Length.mm.) + geom_point()
p2.males1_nosnakes <- p2.males1 %>% filter(Body.depth.mm.>80)

snakes <- p2.males1 %>% filter(Body.depth.mm.<80)
snakes$ID
#doen't look to be too much of a linear difference

#remove 8 and 9, the "snaeky" ones, and see whats up.
#QC 27 and 27, make sure those values are correct

#morpho
p2.lands[,,60]
p2.lands[,,21]


p2.lands.nosnakes<- p2.lands[,,c(-21,-60)]
dim(p2.lands.nosnakes) #good
gpa_nosnakes <- gpagen(p2.lands.nosnakes, curves=pinkcurves) #fuck, it needs sliders
plot(gpa_nosnakes) #looks no diff from orignal
plot(gpa.p2.orig) #original
dim(p2.lands)


pca_nosnakes <- gm.prcomp(gpa_nosnakes$coords)
  
plot(pca_nosnakes)
plot(pca.p2.orig)


#separate RW's for ggplot
df_nosnakes <- as.data.frame(pca_nosnakes$x[,1:3])

wild.or.hatch_nosnakes <- wild.or.hatch[-c(21,60),]

df_nosnakes$origin <- wild.or.hatch_nosnakes$Oto
df_nosnakes$ID <- wild.or.hatch_nosnakes$Fish.ID


ggplot(df_nosnakes) + aes(x=Comp3, y=Comp1, color=origin) + geom_point() + stat_ellipse()
#looks really similar to with snakes

```




Revisit 07/22/22
```{r}
summary(mod.p2021.depth_log) #the important part for analysis
summary(mod.p2021.snout_log)

summary(Snout_global_pinkodd_julian)
summary(Depth_global_pinkodd_julian)
```


